Simple connectome inference from partial correlation statistics in calcium imaging

### 2014-06-23 article.tex
remove comment
explain which noise in pca

At this step, the processed signal only contains clean spikes. However, when a
large part of the network is firing, detecting pairwise neuron interactions is a
difficult task. Furthermore, it is even possible that neurons spike because of
a global high activity, without any actual cause (see \cite{stetter2012model}).
% \textcolor{red}{G: say that neurons may fire because of global
% high activity, without any actual cause -- see generator.}
Accordingly, in order to reduce false interaction detection,
peak intensities at time $t$ are regularized by the global activity of all
neurons, thereby reducing the importance of peaks identified during global
high activity periods:
\begin{align}
 w(x^{t}_i) &= (x^{t}_i + 1 )^{1 + \frac{1}{\sum_{j} x^{t}_j}}.
\end{align}

@@

Under the assumption that random variables follow a joint Gaussian
distribution $\mathcal{N}(\mu; \Sigma)$ of mean vector $\mu$ and covariance
matrix $\Sigma$, $X_i$ and $X_j$ are conditionally independent with respect to
the other variables $X^{-i,j}$ if their \textit{partial correlation} statistic
is null~\citep{koller2009probabilistic}, i.e., if
\begin{equation}
\rho_{X_i, X_j | X^{-i,j}} = -\frac{\Sigma^{-1}_{ij}}{\sqrt{\Sigma^{-1}_{ii} \Sigma^{-1}_{jj}}} = 0.
\end{equation}
Conversely, $X_i$ and $X_j$ are conditionally dependent if $\rho_{X_i, X_j |
X^{-i,j}} \neq 0$, where $\rho_{X_i, X_j | X^{-i,j}}$ can now be interpreted as
the degree of association between the two variables (i.e., the larger the
value, the stronger the interaction). Note however that the partial correlation
statistic is symmetric (i.e., $\rho_{X_i, X_j | X^{-i,j}} = \rho_{X_j, X_i |
X^{-j,i}}$) and therefore does not account for the direction of the
interaction. Together, the partial correlation statistics correspond to the
inverse $\Sigma^{-1}$ of the covariance matrix, also known as the concentration
or precision matrix. Computationally, they can be estimated efficiently in
$O(M^3)$ (instead of $O(p^3)$) from the $M$ first principal components
extracted through PCA~\citep{bishop2006pattern} on the processed data. In our
experiments, $M$ is set to $0.8 p$, thereby allowing us to further filter noise
in the data.

>>>>

At this step, the processed signal only contains clean spikes. However, when a
large part of the network is firing, detecting pairwise neuron interactions is
a difficult task. Furthermore, it is even possible that neurons spike because
of a global high activity, without any actual cause (see
\cite{stetter2012model}). Accordingly, in order to reduce false interaction
detection, peak intensities at time $t$ are regularized by the global activity
of all neurons, thereby reducing the importance of peaks identified during
global high activity periods:
\begin{align}
 w(x^{t}_i) &= (x^{t}_i + 1 )^{1 + \frac{1}{\sum_{j} x^{t}_j}}.
\end{align}

@@

Under the assumption that random variables follow a joint Gaussian
distribution $\mathcal{N}(\mu; \Sigma)$ of mean vector $\mu$ and covariance
matrix $\Sigma$, $X_i$ and $X_j$ are conditionally independent with respect to
the other variables $X^{-i,j}$ if their \textit{partial correlation} statistic
is null~\citep{koller2009probabilistic}, i.e., if
\begin{equation}
\rho_{X_i, X_j | X^{-i,j}} = -\frac{\Sigma^{-1}_{ij}}{\sqrt{\Sigma^{-1}_{ii} \Sigma^{-1}_{jj}}} = 0.
\end{equation}
Conversely, $X_i$ and $X_j$ are conditionally dependent if $\rho_{X_i, X_j |
X^{-i,j}} \neq 0$, where $\rho_{X_i, X_j | X^{-i,j}}$ can now be interpreted as
the degree of association between the two variables (i.e., the larger the
value, the stronger the interaction). Note however that the partial correlation
statistic is symmetric (i.e., $\rho_{X_i, X_j | X^{-i,j}} = \rho_{X_j, X_i |
X^{-j,i}}$) and therefore does not account for the direction of the
interaction. Together, the partial correlation statistics correspond to the
inverse $\Sigma^{-1}$ of the covariance matrix, also known as the concentration
or precision matrix. Computationally, they can be estimated efficiently in
$O(M^3)$ (instead of $O(p^3)$) from the $M$ first principal components
extracted through PCA~\citep{bishop2006pattern} on the processed data. In our
experiments, $M$ is set to $0.8 p$, thereby allowing us to further filter the
the data by removing the contribution of principal components which are
likely due to noise.
